---
title: "Nowcasting in the UK during the COVID-19 pandemic"
author: "Ryan Teo"
format: 
  revealjs:
    toc: true
    toc-depth: 1
    toc-title: "Contents"
    incremental: true
    transition: slide
    transition-speed: fast
    slide-number: c
    chalkboard: true
---

```{r, include = FALSE}
require(tidyverse)
require(here)
require(data.table)
require(ggpubr)
source(here("code", "plot-fns.R"))

# load data
dt <- readRDS(here("data", "observations", "all.rds"))
```

# Introduction

## Nowcasting
- What is nowcasting?
- Why is it important?

## Nowcasting for COVID-19
- Past studies
- Gap

## Objectives
1. Explore the utility of nowcasts of COVID-19 incidence conducted on publicly available data in England
2. Assess the impact of reporting structures on the accuracy of forecasts
    + Weekends
    + Bank holidays
    + Reporting schedules

# Methodology

## Dataset
- New COVID-19 cases by specimen date^[date during which test was taken]
    + Historical data available
    + Available from 23 Feb 2021 onwards
    + Reported every week day up till 1 Jul 2022 and weekly thereafter

## Dataset

```{r, echo = FALSE}
p_case_delay <- plot_case_delay(dt, 7)
```

## Nowcasting
- Expected final counts
- Delay distributions
- Observation model

## Models
1. [Fixed]{style="color: red;"} reporting delay distribution across time
2. Varying reporting delay distribution by [weekends]{style="color: red;"}
3. Varying reporting delay distribution by [day of week]{style="color: red;"} and [bank holidays]{style="color: red;"}
4. Model 3 on [weekly reported]{style="color: red;"} data

## Model evaluation
- Weighted Interval Score

# Results

## Empirical Reporting Delays

## Nowcasting Models

## Scores

## Diagnostics

# Discussion

# Conclusions